Supplemental Material for ``Assessing United States county-level exposure for research on tropical cyclones and human health"

### 2020-09-02 ehp_revision/response_to_reviewers.Rmd

\textit{Lines 294--295 from original manuscript:}

\begin{quotation}
\noindent
\textit{"The tornado observations from this dataset form a traditional tornado
event database for the US, and so we did not further validate the tornado event
data."}
\end{quotation}
\end{shaded}

**Response:**

This database is NOAA's database of records on tornado events in the United
States. It is the original source for the only other national tornado database
available for the country (the Storm Prediction Center [SPC]'s National Tornado
Database), which provides a version of the same data in which tornado events are
joined into a single record in cases where they cross county lines. There is no
other comparable source of data on tornado occurrences that could be used for a
comparison as validation.

The NOAA Storm Events database originated as a database for recording tornadoes
in the US in 1953, and was originally called the Climatological Data National
Summary [@Ashley2007]. The database has gone through a number of expansions to
cover more events, and it now includes reports of other types of
hydrometeorological disaster events, including floods, extreme heat and cold,
wildfires, frost, and hail. However, it has maintained its status as the
database of record for tornadoes in the United States.

We have changed the text in the manuscript to make this clearer:

> "The tornado observations from this dataset---**along with the derived version
of the data available through the Storm Prediction Center's National Tornado
Database---are the tornado event data of record** for the US. **There are no other
collections of tornado data comparable in temporal or geographical scope,** and so
we did not further validate the tornado event data."

### 2020-09-02 ehp_revision/response_to_reviewers.Rmd

1. Added discussion on the question of county-level versus individual-level
data. 
2. Added explanations for how our comparisons among different storm hazards
can help epidemiologists in designing studies and statistical analysis.
3. Added a small analysis of the correlation in rainfall metrics specifically
in cases of high rainfall ($\ge75$ mm of cumulative rainfall in the county
associated with the storm), to deepen the discussion of potential disagreement
in continuous measures of rainfall during extreme conditions. 
4. Changed labeling, axis range, and other presentation details for some
figures, as well as confirmed that the color scales will be accessible.
5. Corrected typos and improved wording throughout based on suggestions of the
reviewers.

# Reviewer 1: {-#reviewer-1:}

## Overall comments {-#overall-comments}
\colorlet{shadecolor}{blue!8}
\begin{shaded}
\textbf{Reviewer 1 Overview:}
I applaud the authors for this work, and their efforts to develop standardized
ways to measure exposure to tropical cyclones. As a disaster scientist, I very
much appreciate efforts that promote our collective ability to learn from and
across disasters to build cumulative science. 
\end{shaded}

**Response:** Thank you for your helpful suggestions. We have responded to 
each in detail below.

@@

While I enjoyed reading the manuscript very much, I have one question I could
not figure out clearly.

One objective is to investigate patterns and agreement between different
exposure measures. I do not clearly understand the value of studying agreement
between exposures, while each exposure should be measured separately, so both
individual and overall effects can be assessed.  I do understand using "distance
from the storm" as exposure is only proxy measure and highly sensitive to
misclassification, but because direct storm hazards such as rainfall, flood,
storm surge, and tornado are publicly accessible as authors indicated, authors
could just simply guide readers to use the best exposure data rather than
conducting a quite intense analysis. This may be due to my limited knowledge, so
it would be appreciated if the authors provide explicit explanations on the
purpose and benefit of the analysis.

\end{shaded}

Thank you for this suggestion, and we can see that we could have more clearly 
articulated in the original manuscript how our assessment of patterns in and 
among exposures to specific hazards can help epidemiologists to design their
studies and the statistical methods to analyze the resulting data. We provide
a detailed response in our response below to Comment 1 from this reviewer.

We have moved other suggestions that you sent in "Track Changes" of the Word 
document for the original manuscript and address each in detail below. When 
there were several comments on a related topic, we have grouped them together
to provide a more cohesive response to the point.

\begin{shaded}
\textbf{Reviewer 2 (R2) Comment 1:}

Several of the comments added by the reviewer in the Word document focused on the
purpose of calculating agreement among hazard-based metrics. These were:

\begin{itemize}
\item Line 164: Isn't it obvious to use better exposure measures than proxies?
Wind, rainfall, tornado, and flood are direct products of storms while distance
is only proxy of those exposures. Rather than calculating agreement among the
exposures, it would be important to note that these are independent hazards that
can come together or come with different intensity and at different time
windows.
\item Line 354: I am not sure why different exposures should be compared and
measured the agreement. I would just treat each exposure as single independent
exposure and give a summary score by summing all the exposure with some
weighting for multiple exposures. I may not understand the importance of this
approach. If so, adding the explanation would help readers not to questioning.
\item Subsection of Results on \textit{Agreement across exposure metrics}: 
As stated below, storms bring different hazards to different locations with
different time. What would be new learning by comparing these different pairs of
exposure metrics. Somewhat apple and orange comparison. Maybe providing more
explanation can help what the important learning from this is. I may not catch
it. Similar comment was added earlier.
\item Lines 494--495: Why this agreement is important? Same questions added above
already, but still wonder what I missed. It seems like the authors try to
develop an exposure matrix that is common in occupational epidemiology, but it
is only valid method when the exposures are similar and lead to the same health
outcomes.  Rain, wind, flood, and tornado should be treated separately, maybe
except rain and flood. It is possible to be an extreme storm with only severe
wind without major impacts from other exposures. Tornado is a good example.
\end{itemize}

The text in the manuscript for comments with line numbers are reproduced below:

\begin{quotation}
\noindent
\textit{"They found important differences, concluding that a study may be prone
to bias from exposure misclassification if distance to the storm track is used
as a proxy for exposure (Grabich et al. 2015a)."}
\end{quotation}

\begin{quotation}
\noindent
\textit{"Finally, we investigated how well exposure assessment agreed across
these metrics."}
\end{quotation}

\begin{quotation}
\noindent
\textit{"For another set of tropical cyclones (e.g., Ernesto in 2006, and Bertha
in 1996, Isabel in 2003), there was moderate to good agreement for pairwise
combinations of distance, rain, and wind, but poor agreement for other
combinations of metrics, while for another set of storms (e.g., Matthew in 2004
and Katrina in 2005), there was moderate to good agreement between distance and
rain."}
\end{quotation}

@@

If measurements of one storm hazard are used as a proxy to capture exposure to
a different hazard, or a non-hazard proxy like distance to the storm track is
used to assess exposure to storm hazards, the resulting exposure
misclassification could be differential (i.e., associated with the outcome of
interest or with factors associated with risk of the outcome of interest).

Therefore for most tropical cyclones, since exposure assessments often differ
across storm hazards, as well as between each storm hazard and a distance-based
proxy measurement, exposure misclassification is a potential risk.

Further, we investigated agreement in exposure assessment across these metrics
when applied individually. We found large differences in which counties are
exposed to different hazards of tropical cyclones and that distance is, at
best, a moderate, and often a very poor, surrogate for exposure to the specific
tropical cyclone hazards of high wind, extreme rainfall, flooding, and
tornadoes. Use of distance as a surrogate for any of these hazards, or exposure
assessment based on one hazard when the pathway for health impacts is in part
or full through another storm hazard, could lead to exposure misclassification.
In the case of tropical cyclone risk and impact studies, including
epidemiological studies, this would result in biased estimates.  Our findings
highlight the importance of clarifying the potential pathway from tropical
cyclone hazards to health impacts when conducting tropical cyclone
epidemiological studies, and then basing exposure assessment on these hazards.

---------------------------------------------------------------------------

@@

> "To develop efficient public health interventions addressing the obesity 
epidemic, it is important to identify exactly which aspects of the
environment influence obesity risk. For example, demonstrating that the density 
of fast-food restaurants was associated with obesity risk and demnstrating that
the density of sports facilities was associated with obesity risk would lead to 
different interventions." [@leal2012multicollinearity]

However, it can be very difficult to separate the effects of different factors
if those factors are strongly correlated [@leal2012multicollinearity]:

> "However, many neighborhood characteristics, especially those related to the
densities of physical features and services, are strongly correlated with each
other. Therefore, it is not clear whether it is even possible to disentangle the
effects of these different environmental dimensions, even if they are
hypothesized to influence obesity risk through distinct causal pathways."
[@leal2012multicollinearity]

-----------------------------------------------------------------------------

@@

However, if our analysis had found that hazard exposures follow very similar
patterns, in terms of the counties exposed during a storm, then this strong
agreement would result in a number of problems in fitting and interpreting these
types of models. First, strong agreement in two or more of the explanatory
variables in a regression model leads to collinearity, which among other
implications often leads to a large inflation of confidence intervals for the
effect estimates of the correlated explanatory variables. Further, if the hazard
exposures were in strong agreement, then it would be difficult or impossible to
disentagle the effects of the separate hazards, as a strong effect for one would
be impossible to interpret with confidence, since the strong agreement in the
data (and resulting low co-variability in the variables in the data) could cause
the model to mistakenly attribute the effect of one hazard to the other.
Further, it may become impossible in this case to estimate synergistic effects
between the two, as an indicator of an interaction between the two exposures
would be almost collinear with the two main effects.

In our analysis, we find that this is usually not the case for the four tropical
cyclone hazards that we consider. Instead, we find that for most storms, the
hazards are fairly well-separated in terms of the counties they effect. There is
a particularly strong pattern where inland counties can be exposed to extreme
rain and flooding, but not wind. This separation opens the door for some
interesting study designs and statistical modeling. For example, these patterns
suggest that a study could investigate whether storm-associated rains alone
create a health risk, as there are many exposures with only rain in inland
counties, and so health associations could be explored here and contrasted with
coastal areas that experienced both wind and extreme rain.

-------------------------------------------------------------------------------

\begin{shaded}
\textbf{R2 Comment 2:}

Other comments were related to correlations estimated in the validation 
analysis for the precipitation data:

\begin{itemize}
\item Line 394: It looks like the correlation coefficients can be very low if measured
from 75mm or somewhere around.  This seems to be a significant issue for
analyzing storms with heavier rainfall using NLDAS-2 data.
\item Line 515: It would be informative to provide correlation coefficient for only
above 75 mm measures.
\end{itemize}

The text from the referenced lines are reproduced below:

@@

> **Text in original manuscript:**
"The rainfall data are generally well-correlated with ground-based observations,
but may sometimes underestimate very high rainfall values (Figure 2). When
rainfall data is used to create binary exposure classifications, this
disagreement is unlikely to influence results, as both data sources agree in
identifying these as storms with high rainfall, but would be important to
consider for cases that include rainfall as a continuous measurement."

We have added estimates of these correlations specifically for
high-precipitation events in a new table in the Supplemental Material (Table S1,
reproduced as Table \@ref(tab:highprecipcorr) of this response document).

\input{tables/precip_high_corr.tex}

---------------------------------------------------------------------------

This is a great suggestion and helps provided quantitative support to the text
in the paper that the reviewer highlighted when making this comment. We have
added a table in the Supplemental Material that provides the Spearman
correlation coefficients for both all tropical cyclones to come near the sample
counties (replicating the analysis shown in Figure 2 of the main text) and
then when the analysis is limited to only those events with cumulative
precipitation of 75 mm or higher based on either data source.

\begin{shaded}
\textbf{R2 Comment 3:}
Figure 4: I would add Y-axis label "NOAA classification."
\end{shaded}

@@

This is a great suggestion. We have made this change, as well as several changes
in response to EHP's Request 4 later in this response.

The original and revised versions of Figure 4 of the paper are reproduced in
Figure \@ref(fig:fig4) of this response.

```{r fig4, fig.cap="Reproduction of Figure 4 from the original manuscript (top) and revised version in the resubmitted manuscript (bottom). In the revised version, we have added a title to the y-axis ('Classification based on NOAA Storm Events Data'), binned the values shown by each point's fill, so these colors are easier to distinguish, increased the size of the points, and changed the algorithm for jittering, so none of the points overlap.", fig.show = "hold", fig.align = "center", out.width="100%"}
include_graphics("../figures/floodcomparison.pdf")
include_graphics("figures/floodcomparison.pdf")
```

\begin{shaded}
\textbf{R2 Comment 4:}
Table 2: This is another confusing table. Not sure what compared here. The
second column is a mean \# of exposed counties over years and the third column
seems like showing single storm with the max \# of exposed counties. Not sure
what info I should digest from this table.
\end{shaded}

@@

**Response:**

We hope that these results can help epidemiologists who are planning new
multi-year studies of tropical cyclones and associated health risks. For 
planning a multi-year study, it is helpful to know how many exposures the 
study might incorporate, as this can help in giving a general idea of whether
the study might have adequate statistical power to investigate a certain 
research question. This table provides some basic summaries of the extent
of exposure to different tropical cyclone hazards, both on average for the
storms in the study and also for those storms that were most extensive. 
Further, as with other results in the study, it provides evidence that 
patterns in different hazards can vary substantially, both because some 
hazards are much less likely to expose counties (e.g., torndaoes are a fairly 
rare exposure, based on this table) and also because there is variation in 
the most extensive storms across the hazards considered.

We have added text to help the readers interpret this table ...

\begin{shaded}
\textbf{R2 Comment 5:}
Line 602: This is not a new finding, but how much misclassified would be a good
question to answer.

@@

\textit{Line 600--602 from original manuscript:}

\begin{quotation}
\noindent
\textit{"Based on our results, the use of a distance-based metric to assess
exposure to any of these hazards, or the use of measurements from one hazard as
a proxy for exposure to any of the other hazards considered, would often
introduce exposure misclassification."}
\end{quotation}
\end{shaded}

**Response:** 
[Add new analysis of what percent of counties misclassified if distance 
is used as proxy for each hazard?]

\begin{shaded}
\textbf{R2 Comment 6:}
Lines 626--627: It would still be beneficial to treat each exposure as a single
variable first and then consider overall impact. Only if each exposure data are
available.

@@

# Reviewer 3: {-#reviewer-3}

\begin{shaded}
\textbf{Reviewer 3 overview:}
This paper describes the development of a dataset and R package that can be used
to assess exposure to tropical cyclone-related hazards (e.g., wind,
precipitation, flooding) at the county level in the eastern United States, as
well as a descriptive analysis of county-level exposure to these hazards. This R
package is novel and I believe of high interest to researchers interested in the
health effects of tropical cyclones.  I have a few minor suggestions to further
improve the clarity of the manuscript.
\end{shaded}

**Response:**

Thank you for the helpful suggestions and recommendations. We have responded in 
detail to each comment or question below.

@@

\begin{quotation}
\noindent
\textit{"To identify flood- and tornado-based tropical cyclone exposures in US
counties, we matched storm tracks with event listings from the National Oceanic
and Atmospheric Administration (NOAA)’s Storm Event Database (NOAA NCEI 2020).
While this database has recorded storm data, particularly tornadoes, since 1950,
its coverage changed substantially in 1996 to cover more types of storm events,
including flood events (NOAA NCEI 2020). We therefore only considered flood
metrics of tropical cyclone exposure for storms in 1996 and later."}
\end{quotation}
\end{shaded}

**Response:**

As suggested, we have added that these events are based on reports in the
Methods, in addition to our previous discussion of the point in the Discussion:

> **Revised text (relevant addition in bold):**
To identify flood- and tornado-based tropical cyclone exposures in US counties,
we matched storm tracks with event listings from the National Oceanic and
Atmospheric Administration (NOAA)’s Storm Event Database (NOAA NCEI 2020).
**Events are included in this database based on reports from National Weather
Service personnel and other sources.** While this database has recorded storm
data, particularly tornadoes, since 1950, its coverage changed substantially in
1996 to cover more types of storm events, including flood events (NOAA NCEI
2020). We therefore only considered flood metrics of tropical cyclone exposure
for storms in 1996 and later."

@@

\textit{Lines 294--295 from original manuscript:}

\begin{quotation}
\noindent
\textit{"The tornado observations from this dataset form a traditional tornado
event database for the US, and so we did not further validate the tornado event
data."}
\end{quotation}
\end{shaded}

**Response:**

This database is NOAA's official database on tornado events in the United States
[@mccarthy2003nws]. It is the original source for the only other national
tornado database available for the country (the Storm Prediction Center [SPC]'s
National Tornado Database [@center2020storm]), which provides a version of the
same data in which tornado events are joined into a single record in cases where
they cross county lines. There is no other comparable source of data on tornado
occurrences that could be used for a comparison as validation.

@@

The NOAA Storm Events database originated as a database for recording tornadoes
in the US in 1953, and was originally called the Climatological Data National
Summary [@Ashley2007]. The database has gone through a number of expansions to
cover more events, and it now includes reports of other types of
hydrometeorological disaster events, including floods, extreme heat and cold,
wildfires, frost, and hail. However, it has maintained its status as the
database of record for tornadoes in the United States.

We have changed the text in the manuscript to make this clearer, changing the
wording to clarify that this is the official NOAA data on tornado events in the
US and that there are no other data sources of comparable scope that could be
used for validation:

> **Text added to the revised manuscript (relevant additions in bold):**
"The tornado observations from this dataset---**along with the derived version
of the data available through the Storm Prediction Center's National Tornado
Database [@center2020storm]---are the official tornado event data** for the US.
**There are no other collections of tornado data comparable in temporal or
geographical scope,** and so we did not further validate the tornado event
data."

@@

20-km and as large as 200-km [@mallin2006; @quiring2011variations].

\begin{shaded}
\textbf{R3 Comment 7:}
Figure 3: The x-axis refers to one of the two wind data sources as "Extended
Best Tracks." Is this the same thing as the wind radii dataset described in the
text? Is so, I suggest clarifying this in the figure label.
\end{shaded}

**Response:**

Yes, it is the same. We have revised the x-axis on the figure to clarify that
the comparison is based on the wind radii dataset described in the text, as
requested. The original and revised versions of Figure 3 of the manuscript are
reproduced in this response document as Figure \@ref(fig:fig3).

### 2020-09-02 ehp_revision/writing/ehp_methods.tex
Work some more on response

To identify flood- and tornado-based tropical cyclone exposures in \ac{US}
counties, we matched storm tracks with event listings from the National Oceanic
and Atmospheric Administration (NOAA)'s Storm Event
Database~\parencite{stormevents}. \textbf{Events are listed in this database
based on reports from National Weather Service personnel and other sources.}
While this database has recorded storm data, particularly tornadoes,
since~1950, its coverage changed substantially in~1996 to cover more types of
storm events, including flood events~\parencite{stormevents}. We therefore only
considered flood metrics of tropical cyclone exposure for storms in~1996 and
later.

@@

The tornado observations from this dataset form a traditional tornado event
database for the \ac{US}, and so we did not further validate
the tornado event data. It is difficult to characterize flooding at the county
level because flooding can be very localized and can be triggered by a variety
of causes. To investigate the extent to which the NOAA flood event listings
capture extremes that might be identified with other flooding data sources, we
investigated a sample of study counties, comparing the flood event data during
tropical storms with streamflow measurements at \ac{US} Geological Survey
county streamflow gages \parencite{usgsgages, countyfloods, dataRetrieval}.  
We considered nine study counties, selecting counties geographically spread
through storm-prone areas of the eastern \ac{US} and with at least one streamflow
gage reporting data during all events. The sample counties were: Baltimore County,
MD; Bergen County, NJ; Escambia County, FL; Fairfield County, CT; Fulton
County, GA; Harris County, TX; Mobile County, AL; Montgomery County, PA; and
Wake County, NC.

>>>>

To identify flood- and tornado-based tropical cyclone exposures in \ac{US}
counties, we matched storm tracks with event listings from the National Oceanic
and Atmospheric Administration (NOAA)'s Storm Event
Database~\parencite{stormevents}. \textbf{Events are included in this database
based on reports from National Weather Service personnel and other sources.}
While this database has recorded storm data, particularly tornadoes,
since~1950, its coverage changed substantially in~1996 to cover more types of
storm events, including flood events~\parencite{stormevents}. We therefore only
considered flood metrics of tropical cyclone exposure for storms in~1996 and
later.

@@

The tornado observations from this dataset---\textbf{along with the derived
version of the data available through the Storm Prediction Center's National
Tornado Database~\parencite{center2020storm}---are the official tornado event
data} for the \ac{US}. \textbf{There are no other collections of tornado data
comparable in emporal or geographic scope,} and so we did not further validate
the tornado event data. It is difficult to characterize flooding at the county
level because flooding can be very localized and can be triggered by a variety
of causes. To investigate the extent to which the NOAA flood event listings
capture extremes that might be identified with other flooding data sources, we
investigated a sample of study counties, comparing the flood event data during
tropical storms with streamflow measurements at \ac{US} Geological Survey
county streamflow gages \parencite{usgsgages, countyfloods, dataRetrieval}.  We
considered nine study counties, selecting counties geographically spread
through storm-prone areas of the eastern \ac{US} and with at least one
streamflow gage reporting data during all events. The sample counties were:
Baltimore County, MD; Bergen County, NJ; Escambia County, FL; Fairfield County,
CT; Fulton County, GA; Harris County, TX; Mobile County, AL; Montgomery County,
PA; and Wake County, NC.

### 2020-09-02 ehp_revision/writing/ehp_response_darlings.tex
Work some more on response

>>>>

R3 Comment 5

This database serves as the basis for much of the large-scale (e.g., assessing
patterns over the US over long time periods) tornado-based research for the US. 
It does have some well-documented limitations. One is a notable increase over 
time in the number of tornadoes reported, particularly for the weakest, F0 [?]
tornadoes. [This stabilizes some in the age of radar?]

> "Weather events listed in the publication Storm Data are considered to be
the official database for tornadoes by the National Weather Service." 
[written by someone at the NWS Storm Prediction Center] [@mccarthy2003nws]

The US National Weather Service's Storm Prediction Center has a tornado database
called the "National Tornado Database" [@center2020storm]. This uses the NOAA
Storm Events database as its primary source [@center2020storm]. They note on their
website:

> "The tables below provide the links to comma separated values (.csv) files for
tornadoes, hail, and damaging winds, as compiled from *NWS Storm Data*. Tornado 
reports exist back to 1950 while hail and damaging wind events date from 1955.

The NOAA Storm Events database, or the Storm Prediction Center's version of the
tornado database have been used in a number of studies, including in a number of
studies of tornado climatology [@tippett2016more; @brooks2003climatological;
@strader2015climatology].

> "The dataset we will use is the so-called smooth log of severe-weather reports
collected by the SPC and archived in the National Oceanic and Atmospheric 
Administration publication *Storm Data*." [@brooks2003climatological]

> "Data such as tornado magnitude, path length, path width, and so on are
recorded by the National Climatic Data Center (NCDC) as a public service and for
use in meteorological, climatological, and engineering studies (Edwards et al.
2013)." [@strader2015climatology]

Sometimes used in conjunction with other NWS data products:

> "Initially, tornado cases available in the NWS's Damage Assessment Toolkit
(NOAA, 2014) were employed, which is a GIS-based framework for collecting,
storing, and retrieving damage survey data (Camp et al., 2014). This toolkit
provides a variety of tornado event filtering (tornado survey point, track,
footprint, and/or swath) and download options [key mark-up language (.kml) or
shapefile (.shp)], supplying an initial sample of georeferenced damage
assessments. In this study, a tornado footprint is defined as the maximum areal
extent of tornado intensity inferred by the damage, wind speed measured directly
by mobile Doppler radar, or assessed theoretically as the recorded length
multiplied with the maximum width as reported in Storm Data."
[@strader2015climatology]

> "To serve the public interest and the National Climatic Data Center (NCDC)
storm data record, the National Weather Service (NWS) documents the path length,
width, and maximum damage rating for every tornado county segment (NOAA
2007).[Footnote in manuscript: County segments of tornado paths are then
combined at the NWS Storm Prediction Center to yield a unified one-tornado
(ONETOR) dataset of whole-tornado records (Schaefer and Edwards 1999).] Such
data are used in meteorological and climatological research, as well as in
determining construction standard for critical infrastructure such as
high-tension electric lines and nuclear power plants (e.g., Ramsdell et al.
2007)." [@edwards2013tornado]

Sometimes used in conjunction with other datasets:

> "The contiguous U.S. tornado fatality dataset utilized in this study was 
transcribed and compiled from two primary sources: (1) a long-term study of 
U.S. tornados by Grazulis (1993, 1997, hereafter Grazulis dataset) and 
(2) the National Climatic Data Center's Storm Data (NCDC 1959--2005) and 
'Storm Events' datavase. The Grazulis dataset included 'significant' tornado
events, including all events that are known to have produced a fatality, 
from 1680 to 1995. Storm Data reports from 1959 to 2005 were utilized to 
supplement the existing record of killer tornado events documented by 
Frazulis." [in a study of tornado fatalities] [@Ashley2007]

History:

> "During the early 1950s, the US Weather Bureau began a concerted effort to
count all US tornadoes. In 1950, a list of tornadoes was provided in the
*Climatological Data National Summary*, but it was not until 1953, the first
full year of the issuance of Weather Bureau tornado watches, that the agency
began to formally count all tornadoes. The *Climaticalogical Data National
Summary* document evolved into *Storm Data* by 1959, which, in addition to
tornado data, included information on other storm perils, such as severe hail,
high winds, and floods (Grazulis 1993). Over the years, the tornado dataset has
been formatted and adjusted by the National Severe Storms Forecast Center and
Storm Prediction Center (SPC), and currently the National Tornado Database is
administered by the NWS Headquarters, SPC, and NCDC (McCarthy 2003)."
[@Ashley2007]
